---
title: "DASM"
output: html_document
date: "2026-06-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(stringr)
library(ggplot2)
library(dplyr)

if (!requireNamespace("WRS2", quietly = TRUE)) {
  install.packages("WRS2")
}
if (!requireNamespace("janitor", quietly = TRUE)) {
  install.packages("janitor")
}

if (!requireNamespace("patchwork", quietly = TRUE)) {
  install.packages("patchwork")
}
if (!requireNamespace("survey", quietly = TRUE)) {
  install.packages("survey")
}
if (!requireNamespace("car", quietly = TRUE)) {
  install.packages("car")
}

library(janitor)
library(patchwork)
library(car)
```

```{r}

load_orig_dataset <- function(){
  
  all_results = read.csv("Survey_results_2026.csv")
  dim(all_results)
  colnames(all_results)
  
  return (all_results)
}

create_g13_subset <- function(all_results){
  
  df = all_results %>% select(starts_with("DQ") | starts_with("G13"))
  colnames(df) = c("age", "gender", "academic_program", "Q1", "Q2", "Q3")
  
  return(df)
}


show_rundown <- function(df){
  dim(df)
  str(df)
  head(df)
}
```

```{r}
setwd("C:/Users/saram/OneDrive/COURSES/DASM")
all_results <- load_orig_dataset()
df <- create_g13_subset(all_results)
show_rundown(df)
```

### General Exploration before Tackling Research Questions

```{r}
df %>% 
  tabyl(academic_program, gender) %>% 
  adorn_totals(where = c("row", "col"))

gender_over_age_hist <- df %>% 
  count(age, gender) %>% 
  ggplot(aes(x = age, y = n, fill = gender)) + 
  geom_col(color = "black") + 
  labs(title = "Age Distribution by Gender", 
       x = "Age Group", y = "Count", fill = "Gender") + 
  theme_minimal()

prog_over_age_hist <- df %>% 
  count(age, academic_program) %>% 
  ggplot(aes(x = age, y = n, fill = academic_program)) + 
  geom_col(color = "black") + 
  labs(title = "Age Distribution by Academic Program", 
       x = "Age Group", y = "Count", fill = "Program") + 
  theme_minimal()

# Displaying distribution of gender and program over age groups
gender_over_age_hist + prog_over_age_hist
```

We observe that Data Science students span a wider range of age categories 
compared to Business Informatics, where no students from the age groups 18-20 
or over 35 participated in the survey. Data Science has a larger share of 
students with 177 compared to 32 for Business Informatics.
Furthermore, the distribution of female students has a different shape
compared to the overall distribution, as the size of the 21-23 age group
exceeds that of the 23-25 group for females. As male students, 
accounting for 156 out of 209 observations of the sample, appear more in 
23-25 compared to 21-23, they dominate the overall shape too.

## Q1

# After 10 years of successful safety testing, on a scale of 1 to 5 (1 = Strongly Disagree, 5 = Strongly Agree), would you undergo a surgical procedure to implant a cognitive-enhancement chip in your brain? 

```{r}
table(df$Q1)
round(prop.table(table(df$Q1)), 4)
barplot(prop.table(table(df$Q1)), ylim = c(0,0.4))
```

```{r}
df %>% 
  tabyl(academic_program, gender) %>% 
  adorn_totals(where = c("row", "col"))
```


```{r}
library(car)
car::leveneTest(Q1 ~ academic_program * gender, data = df)
```
The homoskedasticity assumption does not hold: traditional lm and anova are not 
reliable. We need to use robust methods to handle this.

```{r}
library(WRS2)
# ==============================================================================
# DIAGNOSTIC STEP: Initial Unweighted Exploration & Robust Checks
# ==============================================================================

# Welch's Alternative to One-Way ANOVA & Pairwise Comparisons
oneway.test(Q1 ~ interaction(academic_program, gender), data = df, var.equal = FALSE)

pairwise.t.test(df$Q1, interaction(df$academic_program, df$gender), 
                p.adjust.method = "holm", pool.sd = FALSE)

# Prepare clean df 
df_robust <- df %>%
  mutate(
    academic_program = as.factor(academic_program),
    gender = as.factor(gender),
    Q1 = as.numeric(Q1) # Mandatory for trimmed means calculation
  )

df_nonparametric <- df %>%
  mutate(
    group_comb = interaction(academic_program, gender),
    Q1_ordered = factor(Q1, levels = 1:5, ordered = TRUE) # Best practice for ranks
  )

# Execute Robust 2-Way ANOVA (Trimming Method)
robust_model <- t2way(Q1 ~ academic_program * gender, data = df_robust)
print(robust_model)

# Execute Non-parametric Multi-group Test
kruskal.test(Q1_ordered ~ group_comb, data = df_nonparametric)
```
We began our analysis by performing a classical Welch’s ANOVA (oneway.test) 
alongside an unweighted pairwise $t$-test with a Holm adjustment across our 
sample groups. This unweighted exploration revealed a single statistically 
significant rift specifically between Data Science Males and Data Science 
Females ($p = 0.030$). Interestingly, this gender divide did not hold between 
Business Informatics Females and Data Science Males, and no main programmatic 
differences were initially detected.

To ensure that these initial unweighted findings were not an artifact of 
breaking the homoskedasticity assumption, we also subjected the raw data to a 
Robust 2-Way Factorial ANOVA via M-measures (WRS2::t2way) and a non-parametric 
Kruskal-Wallis rank test. While both tests flagged significant group-level 
variations driven primarily by gender ($p = 0.044$ and $p = 0.0178$ respectively), 
they suffer from a shared fatal flaw: they cannot accommodate population 
sampling weights.

Relying solely on these unweighted, classical robust and non-parametric 
$p$-values presents severe statistical risks. Our sample suffers from an extreme
structural imbalance, with Data Science students heavily outnumbering Business
Informatics students 177 to 32. Furthermore, Levene’s test explicitly rejects
the assumption of equal variances ($p = 0.0143$), confirming that the 
homoskedasticity assumption is violated. Because classical unweighted tests
treat every respondent equally despite these issues, the overwhelming volume of
the Data Science cohort risks distorting standard errors, altering rank 
distribution, or masking true population-level insights.

Therefore, while this exploratory stage suggests that students share similar 
viewpoints across academic programs, these structural imbalances mean 
traditional descriptive metrics and classical robust regressions are ultimately
inadequate. To rigorously confirm the absence of an academic program effect, 
handle the heteroskedasticity, and validate whether the observed gender divide 
holds true under population adjustments, we must transition to a formal 
multi-factor survey design. By raking the sample to a balanced 50/50 
programmatic split and utilizing robust sandwich variance estimators, we can 
safely finalize these relationships within a design-weighted framework.

```{r}
library(survey)
library(robustbase)
library(stringr)
library(purrr)

# ==============================================================================
# SECTION 1: Define the Survey Design & Weights
# ==============================================================================

# 1. Base design (Equal probability)
base_design <- svydesign(ids = ~1, data = df)

# 2. Rake to adjust for academic program imbalance (50/50 target split)
pop_program <- data.frame(academic_program = c("Data Science", "Business Informatics"), 
                          Freq = c(105.5, 105.5))

weighted_design <- rake(design = base_design,
                        sample.margins = list(~academic_program),
                        population.margins = list(pop_program))

# 3. Clean and prep variables safely within the weighted design framework
weighted_design <- update(weighted_design,
                          Q1_ordered = factor(Q1, levels = 1:5, ordered = TRUE),
                          gender = as.factor(gender),
                          academic_program = as.factor(academic_program))

```

Because Levene's test explicitly rejects the assumption of equal variances 
($p = 0.0143$), traditional linear models and ANOVAs cannot be trusted to 
evaluate these relationships. To test for programmatic effects in the absence of 
homoskedasticity and to account for selection bias, we implement a 
design-weighted survey framework utilizing robust sandwich estimators.

```{r}
# ==============================================================================
# SECTION 2: Initial Multi-Factor Analysis (The "Insignificant" Models)
# ==============================================================================

# 1. Fit a design-weighted linear models
q1_glmodel <- svyglm(Q1 ~ academic_program * gender, design = weighted_design)
summary(q1_glmodel)

q1_ordinal_model <- svyolr(Q1_ordered ~ academic_program * gender, 
                           design = weighted_design)
summary(q1_ordinal_model)

# 2. Run the Design-Corrected Wald Omnibus Test 
# This establishes global interaction significance within the survey framework
# --- Global Omnibus Interaction Test ---
regTermTest(q1_glmodel, ~academic_program:gender)

# 3. Target Subgroup Means Extraction
subgroup_means <- svyby(~Q1, by = ~academic_program + gender, 
                        design = weighted_design, svymean)
print(subgroup_means)

# 4. Target Contrast Isolation (Resolving the baseline masking effect safely)
# --- Specific Target Contrast Test ---
svycontrast(q1_glmodel, contrasts=list("DS_Male_vs_DS_Female" = c(0, 0, 1, 1)))
```

Compared to the default alphabetical reference baseline (Business Informatics
Females) used in the regression outputs, the individual interaction terms do 
not initially appear significant. This masking effect occurs because standard
regression parameters evaluate groups strictly relative to one baseline. 
To verify whether the overall interaction framework holds any statistical
value globally, we perform a design-corrected omnibus test (regTermTest).

In the extracted subgroup means, we notice that the ratings for females 
are lower on average across both academic programs. Given the standard errors, 
we bypass the baseline masking effect by directly contrasting the specific pair
with the largest descriptive gap: Data Science Males versus Data Science Females.

When proper survey controls are applied, the significance of the gender gap
within the Data Science discipline is heavily amplified ($t = 2.927, p \approx 0.003$).
This confirms that the gender rift is a robust, population-level 
phenomenon, while validating that academic programs show no baseline variations.

```{r}
# ==============================================================================
# SECTION 4: Risk Factor Analysis (Q3 Predictors on Q1)
# ==============================================================================

# 1. Parse checklist string dynamically into indicator objects
all_risks <- df$Q3 %>%
  str_split(";") %>%
  flatten_chr() %>%
  unique() %>%
  discard(~ .x == "" | is.na(.x))

q3_dummy_list <- map(all_risks, function(risk) {
  as.numeric(str_detect(df$Q3, fixed(risk)))
})

q3_dummy <- as.data.frame(q3_dummy_list)
colnames(q3_dummy) <- make.names(all_risks)

# Append dynamic columns back to original data stream
df_extended <- cbind(df, q3_dummy)

# 2. Construct a weighted design for the risk dataset
risk_base_design <- svydesign(ids = ~1, data = df_extended)
risk_weighted_design <- rake(design = risk_base_design,
                             sample.margins = list(~academic_program),
                             population.margins = list(pop_program))

# 3. Dynamic Formula Formulation
risk_variables <- colnames(q3_dummy)
formula_string <- paste("Q1 ~", paste(risk_variables, collapse = " + "), "+ gender")
risk_formula <- as.formula(formula_string)

# 4. Run a Design-Weighted Survey GLM for Risk Analysis
# This simultaneously accounts for risk perceptions, sample weights, and heteroskedasticity
survey_risk_model <- svyglm(risk_formula, design = risk_weighted_design)
summary(survey_risk_model)

# 5. Type-II Design-Adjusted Wald Test for Risk Attributes
message("--- Design-Adjusted Type-II Analysis of Deviance ---")
car::Anova(survey_risk_model, test.statistic = "Chisq")
```

## Q2

# After 10 years of successful safety testing, what ratio of the general population (0-100%) do you believe would choose to get the implant?

```{r}
# ==============================================================================
# SECTION 1: Exploratory Data Analysis & Descriptive Inference
# ==============================================================================

# 1. Grouped Boxplot for Visual Exploration
ggplot(df, aes(x = academic_program, y = Q2, fill = gender)) +
  geom_boxplot() +
  labs(title = "Estimated Public Adoption by Program and Gender",
       x = "Academic Program",
       y = "Estimated Population Ratio (0-100%)",
       fill = "Gender") +
  theme_minimal()

# 2. Descriptive Summary Statistics Table
q2_summary <- df %>%
  group_by(academic_program, gender) %>%
  summarise(
    Count = n(),
    Mean = round(mean(Q2, na.rm = TRUE), 2),
    Median = median(Q2, na.rm = TRUE),
    SD = round(sd(Q2, na.rm = TRUE), 2),
    Min = min(Q2, na.rm = TRUE),
    Max = max(Q2, na.rm = TRUE),
    .groups = 'drop'
  )

print(q2_summary)
```

For our exploratory data analysis of Question 2, we generated a grouped
boxplot breaking down public adoption estimates by both academic program
and gender. The summary statistics reveal a shared median estimate of
roughly 30% across most subgroups, though the standard deviations are
exceedingly high (ranging from 21 to 27), indicating a massive degree of
individual uncertainty regarding public sentiment.

Visually, the boxplot exposes a fascinating demographic intersection. In
Business Informatics, male students generally predicted higher adoption
rates than female students, featuring a higher median and an extended
upper quartile. However, in Data Science, this dynamic completely flips.
Female Data Science students hold the most optimistic views of public
adoption, with a median estimate near 45%. Conversely, Data Science
males shift downward to the lowest median in the entire sample. This
visual crossing pattern strongly suggests an interaction effect between
program and gender.

```{r}
# ==============================================================================
# SECTION 2: Analytic Inference (Survey-Weighted Interaction Model)
# ==============================================================================

# Fit a design-weighted linear model for Q2 using the previously established 
# 'weighted_design' to account for the sample imbalance and heteroskedasticity.
q2_interaction_model <- svyglm(Q2 ~ academic_program * gender, design = weighted_design)

# Output the robust summary 
summary(q2_interaction_model)
```

Because Levene's test from Q1 confirmed that the homoskedasticity
assumption is violated across our demographic groups, and given the
extreme 177-to-32 sample imbalance between programs, traditional
unweighted tests like a standard Two-Way ANOVA would yield unreliable
standard errors. To safely test if the visual crossing pattern observed
in the EDA is statistically meaningful, we applied the same
design-weighted generalized linear model (svyglm) utilized in Q1,
seamlessly incorporating our population raking weights.The regression
output confirms the structural dynamics hinted at by our exploratory
visuals. The independent main effects for both academic program
($p = 0.126$) and gender ($p = 0.496$) fail to reach statistical
significance on their own. This indicates that neither demographic trait
acts as a standalone predictor for how a student estimates public
technology adoption.

However, the interaction term between the Data
Science program and Male gender yields a $p$-value of $0.053$. While
sitting precisely on the margin of the traditional $\alpha = 0.05$
threshold, this indicates a strong, marginally significant interaction
trend at the population level. It mathematically validates that the
relationship between gender and adoption estimates is not static; it
dynamically reverses depending on the student's academic specialization.
Therefore, we conclude that perceptions of societal tech adoption are
not driven by single traits, but are jointly constructed by the
intersection of a student's gender identity and their academic
discipline.

