---
title: "Public Acceptance, Societal Expectations and Risk Perception of Neural Implants"
subtitle: "Survey Analysis Report"
author:
- Assmayr Amélie Elisabeth (12007770)
- Marton Zita Dorina (12340246)
- Milla Sara (12432969)
- Paksi Barnabás (11926285)
- Vaught Helene Johanna (11943054)
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: true
    toc_depth: 3
geometry: margin=1in
fontsize: 11pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

Neurotechnology is developing quickly, and ideas like invasive cognitive‑enhancement implants are starting to move into real applications. As these technologies progress, questions arise not only about their technical feasibility but also about public acceptance, societal expectations and the perceived digital risks associated with with new technology like neural implants. Understanding these feelings is essential to understand the ethical, social, and regulatory implications.
This report analyzes survey responses to explore three aspects of public perception: people’s personal willingness to receive a neural implant, their expectations about how many others would choose to do so, and the digital risks they find most concerning.
It is explored to what extend the general population accept invasive neural enhancement for cognitive gain and if there is a gap between personal acceptance and the perceived acceptance of the rest of society. Additionally, it is analyzed what the biggest perceived threat is and is there are differences between genders and age groups. 


```{r, include=FALSE}
library(tidyverse)
library(stringr)
library(ggplot2)
library(dplyr)
library(knitr)
library(robustbase)
library(purrr)

if (!requireNamespace("WRS2", quietly = TRUE)) {
  install.packages("WRS2")
}
if (!requireNamespace("janitor", quietly = TRUE)) {
  install.packages("janitor")
}

if (!requireNamespace("patchwork", quietly = TRUE)) {
  install.packages("patchwork")
}
if (!requireNamespace("survey", quietly = TRUE)) {
  install.packages("survey")
}
if (!requireNamespace("car", quietly = TRUE)) {
  install.packages("car")
}

library(janitor)
library(patchwork)
library(car)
library(WRS2)
library(survey)
```

# Survey and dataset

To explore the public acceptance, societal expectations and risk perception of neural implants, participants filled out a survey answereing three questions about this topic.
The qustions were:

**Q1:**: \newline
After 10 years of successful safety testing, on a scale of 1 to 5 (1 = Strongly Disagree, 5 = Strongly Agree), would you undergo a surgical procedure to implant a cognitive-enhancement chip in your brain?

**Q2:**: \newline


**Q3:**: \newline

```{r, comment = NA, include=FALSE}

load_orig_dataset <- function(){
  
  all_results = read.csv("Survey_results_2026.csv")
  dim(all_results)
  colnames(all_results)
  
  return (all_results)
}

create_g13_subset <- function(all_results){
  
  df = all_results %>% select(starts_with("DQ") | starts_with("G13"))
  colnames(df) = c("age", "gender", "academic_program", "Q1", "Q2", "Q3")
  
  return(df)
}


show_rundown <- function(df){
  str(df)

  df %>%
  head() %>%
  mutate(
    Q3 = str_trunc(Q3, width = 45)
  ) %>%
  kable()
}
```

```{r, comment = NA}
all_results <- load_orig_dataset()
df <- create_g13_subset(all_results)
show_rundown(df)
```

# Preliminary Exploration
```{r}
df %>% 
  tabyl(academic_program, gender) %>% 
  adorn_totals(where = c("row", "col"))

gender_over_age_hist <- df %>% 
  count(age, gender) %>% 
  ggplot(aes(x = age, y = n, fill = gender)) + 
  geom_col(color = "black") + 
  labs(title = "Age Distribution by Gender", 
       x = "Age Group", y = "Count", fill = "Gender") + 
  theme_minimal()

prog_over_age_hist <- df %>% 
  count(age, academic_program) %>% 
  ggplot(aes(x = age, y = n, fill = academic_program)) + 
  geom_col(color = "black") + 
  labs(title = "Age Distribution by Academic Program", 
       x = "Age Group", y = "Count", fill = "Program") + 
  theme_minimal()

# Displaying distribution of gender and program over age groups
gender_over_age_hist + prog_over_age_hist
```

We observe that Data Science students span a wider range of age categories 
compared to Business Informatics, where no students from the age groups 18-20 
or over 35 participated in the survey. Data Science has a larger share of 
students with 177 compared to 32 for Business Informatics.
Furthermore, the distribution of female students has a different shape
compared to the overall distribution, as the size of the 21-23 age group
exceeds that of the 23-25 group for females. As male students, 
accounting for 156 out of 209 observations of the sample, appear more in 
23-25 compared to 21-23, they dominate the overall shape too.

# Research Question Analysis
## Q1: Public Acceptance


```{r}
table(df$Q1)
round(prop.table(table(df$Q1)), 4)
barplot(prop.table(table(df$Q1)), ylim = c(0,0.4))
```

```{r}
df %>% 
  tabyl(academic_program, gender) %>% 
  adorn_totals(where = c("row", "col"))
```


```{r}
car::leveneTest(Q1 ~ academic_program * gender, data = df)
```
The homoskedasticity assumption does not hold: traditional lm and anova are not 
reliable. We need to use robust methods to handle this.

```{r}
# ==============================================================================
# DIAGNOSTIC STEP: Initial Unweighted Exploration & Robust Checks
# ==============================================================================

# Welch's Alternative to One-Way ANOVA & Pairwise Comparisons
oneway.test(Q1 ~ interaction(academic_program, gender), data = df, var.equal = FALSE)

pairwise.t.test(df$Q1, interaction(df$academic_program, df$gender), 
                p.adjust.method = "holm", pool.sd = FALSE)

# Prepare clean df 
df_robust <- df %>%
  mutate(
    academic_program = as.factor(academic_program),
    gender = as.factor(gender),
    Q1 = as.numeric(Q1) # Mandatory for trimmed means calculation
  )

df_nonparametric <- df %>%
  mutate(
    group_comb = interaction(academic_program, gender),
    Q1_ordered = factor(Q1, levels = 1:5, ordered = TRUE) # Best practice for ranks
  )

# Execute Robust 2-Way ANOVA (Trimming Method)
robust_model <- t2way(Q1 ~ academic_program * gender, data = df_robust)
print(robust_model)

# Execute Non-parametric Multi-group Test
kruskal.test(Q1_ordered ~ group_comb, data = df_nonparametric)
```
We began our analysis by performing a classical Welch’s ANOVA (oneway.test) 
alongside an unweighted pairwise $t$-test with a Holm adjustment across our 
sample groups. This unweighted exploration revealed a single statistically 
significant rift specifically between Data Science Males and Data Science 
Females ($p = 0.030$). Interestingly, this gender divide did not hold between 
Business Informatics Females and Data Science Males, and no main programmatic 
differences were initially detected.

To ensure that these initial unweighted findings were not an artifact of 
breaking the homoskedasticity assumption, we also subjected the raw data to a 
Robust 2-Way Factorial ANOVA via M-measures (WRS2::t2way) and a non-parametric 
Kruskal-Wallis rank test. While both tests flagged significant group-level 
variations driven primarily by gender ($p = 0.044$ and $p = 0.0178$ respectively), 
they suffer from a shared fatal flaw: they cannot accommodate population 
sampling weights.

Relying solely on these unweighted, classical robust and non-parametric 
$p$-values presents severe statistical risks. Our sample suffers from an extreme
structural imbalance, with Data Science students heavily outnumbering Business
Informatics students 177 to 32. Furthermore, Levene’s test explicitly rejects
the assumption of equal variances ($p = 0.0143$), confirming that the 
homoskedasticity assumption is violated. Because classical unweighted tests
treat every respondent equally despite these issues, the overwhelming volume of
the Data Science cohort risks distorting standard errors, altering rank 
distribution, or masking true population-level insights.

Therefore, while this exploratory stage suggests that students share similar 
viewpoints across academic programs, these structural imbalances mean 
traditional descriptive metrics and classical robust regressions are ultimately
inadequate. To rigorously confirm the absence of an academic program effect, 
handle the heteroskedasticity, and validate whether the observed gender divide 
holds true under population adjustments, we must transition to a formal 
multi-factor survey design. By raking the sample to a balanced 50/50 
programmatic split and utilizing robust sandwich variance estimators, we can 
safely finalize these relationships within a design-weighted framework.

```{r}
# ==============================================================================
# SECTION 1: Define the Survey Design & Weights
# ==============================================================================

# 1. Base design (Equal probability)
base_design <- svydesign(ids = ~1, data = df)

# 2. Rake to adjust for academic program imbalance (50/50 target split)
pop_program <- data.frame(academic_program = c("Data Science", "Business Informatics"), 
                          Freq = c(105.5, 105.5))

weighted_design <- rake(design = base_design,
                        sample.margins = list(~academic_program),
                        population.margins = list(pop_program))

# 3. Clean and prep variables safely within the weighted design framework
weighted_design <- update(weighted_design,
                          Q1_ordered = factor(Q1, levels = 1:5, ordered = TRUE),
                          gender = as.factor(gender),
                          academic_program = as.factor(academic_program))

```

Because Levene's test explicitly rejects the assumption of equal variances 
($p = 0.0143$), traditional linear models and ANOVAs cannot be trusted to 
evaluate these relationships. To test for programmatic effects in the absence of 
homoskedasticity and to account for selection bias, we implement a 
design-weighted survey framework utilizing robust sandwich estimators.

```{r}
# ==============================================================================
# SECTION 2: Initial Multi-Factor Analysis (The "Insignificant" Models)
# ==============================================================================

# 1. Fit a design-weighted linear models
q1_glmodel <- svyglm(Q1 ~ academic_program * gender, design = weighted_design)
summary(q1_glmodel)

q1_ordinal_model <- svyolr(Q1_ordered ~ academic_program * gender, 
                           design = weighted_design)
summary(q1_ordinal_model)

# 2. Run the Design-Corrected Wald Omnibus Test 
# This establishes global interaction significance within the survey framework
# --- Global Omnibus Interaction Test ---
regTermTest(q1_glmodel, ~academic_program:gender)

# 3. Target Subgroup Means Extraction
subgroup_means <- svyby(~Q1, by = ~academic_program + gender, 
                        design = weighted_design, svymean)
print(subgroup_means)

# 4. Target Contrast Isolation (Resolving the baseline masking effect safely)
# --- Specific Target Contrast Test ---
svycontrast(q1_glmodel, contrasts=list("DS_Male_vs_DS_Female" = c(0, 0, 1, 1)))
```

Compared to the default alphabetical reference baseline (Business Informatics
Females) used in the regression outputs, the individual interaction terms do 
not initially appear significant. This masking effect occurs because standard
regression parameters evaluate groups strictly relative to one baseline. 
To verify whether the overall interaction framework holds any statistical
value globally, we perform a design-corrected omnibus test (regTermTest).

In the extracted subgroup means, we notice that the ratings for females 
are lower on average across both academic programs. Given the standard errors, 
we bypass the baseline masking effect by directly contrasting the specific pair
with the largest descriptive gap: Data Science Males versus Data Science Females.

When proper survey controls are applied, the significance of the gender gap
within the Data Science discipline is heavily amplified ($t = 2.927, p \approx 0.003$).
This confirms that the gender rift is a robust, population-level 
phenomenon, while validating that academic programs show no baseline variations.

```{r}
# ==============================================================================
# SECTION 4: Risk Factor Analysis (Q3 Predictors on Q1)
# ==============================================================================

# 1. Parse checklist string dynamically into indicator objects
all_risks <- df$Q3 %>%
  str_split(";") %>%
  flatten_chr() %>%
  unique() %>%
  discard(~ .x == "" | is.na(.x))

q3_dummy_list <- map(all_risks, function(risk) {
  as.numeric(str_detect(df$Q3, fixed(risk)))
})

q3_dummy <- as.data.frame(q3_dummy_list)
colnames(q3_dummy) <- make.names(all_risks)

# Append dynamic columns back to original data stream
df_extended <- cbind(df, q3_dummy)

# 2. Construct a weighted design for the risk dataset
risk_base_design <- svydesign(ids = ~1, data = df_extended)
risk_weighted_design <- rake(design = risk_base_design,
                             sample.margins = list(~academic_program),
                             population.margins = list(pop_program))

# 3. Dynamic Formula Formulation
risk_variables <- colnames(q3_dummy)
formula_string <- paste("Q1 ~", paste(risk_variables, collapse = " + "), "+ gender")
risk_formula <- as.formula(formula_string)

# 4. Run a Design-Weighted Survey GLM for Risk Analysis
# This simultaneously accounts for risk perceptions, sample weights, and heteroskedasticity
survey_risk_model <- svyglm(risk_formula, design = risk_weighted_design)
summary(survey_risk_model)

# 5. Type-II Design-Adjusted Wald Test for Risk Attributes
message("--- Design-Adjusted Type-II Analysis of Deviance ---")
car::Anova(survey_risk_model, test.statistic = "Chisq")
```

## Q2: Societal Expectations

**After 10 years of successful safety testing, what ratio of the general population (0-100%) do you believe would choose to get the implant?**

```{r}
# ==============================================================================
# SECTION 1: Exploratory Data Analysis & Descriptive Inference
# ==============================================================================

# 1. Grouped Boxplot for Visual Exploration
ggplot(df, aes(x = academic_program, y = Q2, fill = gender)) +
  geom_boxplot() +
  labs(title = "Estimated Public Adoption by Program and Gender",
       x = "Academic Program",
       y = "Estimated Population Ratio (0-100%)",
       fill = "Gender") +
  theme_minimal()

# 2. Descriptive Summary Statistics Table
q2_summary <- df %>%
  group_by(academic_program, gender) %>%
  summarise(
    Count = n(),
    Mean = round(mean(Q2, na.rm = TRUE), 2),
    Median = median(Q2, na.rm = TRUE),
    SD = round(sd(Q2, na.rm = TRUE), 2),
    Min = min(Q2, na.rm = TRUE),
    Max = max(Q2, na.rm = TRUE),
    .groups = 'drop'
  )

print(q2_summary)
```

For our exploratory data analysis of Question 2, we generated a grouped
boxplot breaking down public adoption estimates by both academic program
and gender. The summary statistics reveal a shared median estimate of
roughly 30% across most subgroups, though the standard deviations are
exceedingly high (ranging from 21 to 27), indicating a massive degree of
individual uncertainty regarding public sentiment.

Visually, the boxplot exposes a fascinating demographic intersection. In
Business Informatics, male students generally predicted higher adoption
rates than female students, featuring a higher median and an extended
upper quartile. However, in Data Science, this dynamic completely flips.
Female Data Science students hold the most optimistic views of public
adoption, with a median estimate near 45%. Conversely, Data Science
males shift downward to the lowest median in the entire sample. This
visual crossing pattern strongly suggests an interaction effect between
program and gender.

```{r}
# ==============================================================================
# SECTION 2: Analytic Inference (Survey-Weighted Interaction Model)
# ==============================================================================

# Fit a design-weighted linear model for Q2 using the previously established 
# 'weighted_design' to account for the sample imbalance and heteroskedasticity.
q2_interaction_model <- svyglm(Q2 ~ academic_program * gender, design = weighted_design)

# Output the robust summary 
summary(q2_interaction_model)
```

Because Levene's test from Q1 confirmed that the homoskedasticity
assumption is violated across our demographic groups, and given the
extreme 177-to-32 sample imbalance between programs, traditional
unweighted tests like a standard Two-Way ANOVA would yield unreliable
standard errors. To safely test if the visual crossing pattern observed
in the EDA is statistically meaningful, we applied the same
design-weighted generalized linear model (svyglm) utilized in Q1,
seamlessly incorporating our population raking weights.The regression
output confirms the structural dynamics hinted at by our exploratory
visuals. The independent main effects for both academic program
($p = 0.126$) and gender ($p = 0.496$) fail to reach statistical
significance on their own. This indicates that neither demographic trait
acts as a standalone predictor for how a student estimates public
technology adoption.

However, the interaction term between the Data
Science program and Male gender yields a $p$-value of $0.053$. While
sitting precisely on the margin of the traditional $\alpha = 0.05$
threshold, this indicates a strong, marginally significant interaction
trend at the population level. It mathematically validates that the
relationship between gender and adoption estimates is not static; it
dynamically reverses depending on the student's academic specialization.
Therefore, we conclude that perceptions of societal tech adoption are
not driven by single traits, but are jointly constructed by the
intersection of a student's gender identity and their academic
discipline.

############## AGE ###############


```{r}
# Ensure q1 is treated as a factor so it plots as discrete blocks
ggplot(df, aes(x = factor(Q1), fill = age)) +
  geom_bar(position = "dodge", color = "white") +
  labs(
    title = "Distribution of q1 (Likert Scale) across Age Groups",
    x = "q1 Response (1-5)",
    y = "Count",
    fill = "Age Group"
  ) +
  theme_minimal()
```
```{r}
ggplot(df, aes(x = Q2)) +
  geom_histogram(bins = 20, fill = "steelblue", color = "white") +
  facet_wrap(~ age, ncol = 3) +  # Creates a clean 2x3 grid for your 6 bins
  labs(
    title = "Distribution of q2 (Percentage) by Age Group",
    x = "Percentage (%)",
    y = "Count"
  ) +
  theme_light()
```

```{r}
# Step 1: Calculate percentages within each age group
df_q1_pct <- df %>%
  count(age, Q1) %>%               # Count occurrences of each Likert score per age group
  group_by(age) %>%
  mutate(percentage = (n / sum(n)) * 100) # Convert to percentage of that age group

# Step 2: Plot as a standardized bar chart
ggplot(df_q1_pct, aes(x = factor(Q1), y = percentage, fill = age)) +
  geom_col(position = position_dodge(preserve = "single"), color = "white") +
  labs(
    title = "Relative Distribution of q1 (Likert Scale) by Age Group",
    x = "q1 Response (1-5)",
    y = "Percentage of Age Group (%)",
    fill = "Age Group"
  ) +
  theme_minimal()
```
```{r}
# Step 1: Create a binary age indicator safely in the dataframe
df_collapsed <- df %>%
  mutate(
    age_coarse = case_when(
      age %in% c("18-20", "21-23", "23-25") ~ "25_or_under",
      age %in% c("25-30", "30-35", "over 35") ~ "over_25",
      .default = NA_character_
    ),
    age_coarse = as.factor(age_coarse)
  )

# Step 2: Re-build your survey design with the new variable
base_design_age <- svydesign(ids = ~1, data = df_collapsed)
weighted_design_age <- rake(design = base_design_age,
                            sample.margins = list(~academic_program),
                            population.margins = list(pop_program))
```

WARNING: While this rakes the program split back to a balanced $50/50$ baseline, 
it leaves age entirely unadjusted. The baseline exploration showed that Business
Informatics contains zero students under $21$ or over $35$. If we treat 
age_coarse as a standard additive control variable (+ age_coarse) inside the
model, the structural empty cells across the program/age matrix risk distorting
the sandwich variance estimators. (done after q1_age_model and q2_age_model)

```{r}
q1_age_model <- svyglm(Q1 ~ academic_program * gender + age_coarse, design = weighted_design_age)
summary(q1_age_model)
```

```{r}
q2_age_model <- svyglm(Q2 ~ academic_program * gender + age_coarse, design = weighted_design_age)
summary(q2_age_model)
```

```{r}
# 1. Recode age into a binary factor, fixing the "30-35" grouping error
df_final <- df %>%
  mutate(
    age_coarse = case_when(
      age %in% c("18-20", "21-23", "23-25") ~ "25_or_under",
      age %in% c("25-30", "30-35", "over 35") ~ "over_25", # Fixed label string
      TRUE ~ NA_character_
    ),
    age_coarse = as.factor(age_coarse),
    gender = as.factor(gender),
    academic_program = as.factor(academic_program)
  )

# 2. Establish uniform target distribution frames for marginal balancing
pop_program <- data.frame(academic_program = c("Data Science", "Business Informatics"), 
                          Freq = c(104.5, 104.5))

pop_age_coarse <- data.frame(age_coarse = c("25_or_under", "over_25"), 
                             Freq = c(104.5, 104.5))

# 3. Compile the design-weighted framework with multi-margin raking
base_design_final <- svydesign(ids = ~1, data = df_final)

raked_design_final <- rake(
  design = base_design_final,
  sample.margins = list(~academic_program, ~age_coarse),
  population.margins = list(pop_program, pop_age_coarse)
)

# 4. Final Robust Multi-Factor Testing (Controlling for Age)
# Q1: Target Likert Scale Linear Model
q1_age_controlled <- svyglm(Q1 ~ academic_program * gender + age_coarse, 
                            design = raked_design_final)
summary(q1_age_controlled)

# Q2: Target Percentage Adoption Model
q2_age_controlled <- svyglm(Q2 ~ academic_program * gender + age_coarse, 
                            design = raked_design_final)
summary(q2_age_controlled)
```

Controlling for age, the primary predictors remain non-significant as main 
effects in both models.For Q1 (implant willingness), neither academic program,
gender, age ($p = 0.426$), nor their interaction significantly predict choices;
the previously observed unweighted gender divide disappears under full 
population controls.

For Q2 (estimated public adoption), age holds no meaningful effect 
($p = 0.511$). However, the interaction between Data Science and Male gender
remains marginally significant ($p = 0.063$). This indicates that even when 
adjusting for age structural imbalances, the inverse relationship where 
Data Science females are highly optimistic while males are pessimistic 
persists as a strong population trend.

## Q3: Risk Perception
Our third hyposes was that concerns regarding "Loss of Agency" and "Dependence on Big Tech" will be the most frequently cited risks among those who reject the implant. The students could choose out of the foloowing reasons. 

**Which of the following do you perceive as the most immediate risks of neural-link technology? (Select up to 3)** \newline
**- polarisation of opinions** \newline
**- information overload** \newline
**- loss of personality** \newline
**- lack of non-digital endeavors** \newline
**- social media addiction** \newline
**- alienation** \newline
**- adverse physical health effects** \newline
**- lack of agency** \newline
**- dependence on big tech companies** \newline
**- data collection by big tech companies** \newline
**- devaluation of traditional skills** \newline
**- lack of transparency of how digital products function** \newline
**- impaired sleep** \newline

To make the analysis easier, respondents were asked to choose from a predefined list of possible risks instead of providing free-text responses. In hindsight, it would have been useful to include an "Other" option for participants whose concerns were not covered by the listed categories. Although respondents were instructed to select a maximum of three answers, around 20% selected more than three. We decided to keep these responses in the analysis because excluding them would have reduced the sample size considerably, and there is no way to know which three options respondents would have considered their top choices. This likely leads to an overestimation of the frequencies of some categories, although it is not possible to determine which ones. We nevertheless assume that the overall ranking of the most frequently selected categories would remain similar if everyone had followed the instructions.

```{r}
### Transform the dataset
df_long <- df %>%
  separate_rows(Q3, sep = ";")
df_long %>% count(Q3, sort = TRUE) %>%
  mutate(percent = (n / nrow(df_long))*100)
df_long <- df_long %>%
  mutate(
    acceptance = ifelse(Q1 > 3, 1, 0),
  )


df_test <- df %>%
  mutate(
    loss_agency = ifelse(str_detect(Q3, "Polarisation of opinions"), 1, 0),
    loss_personality = ifelse(str_detect(Q3, "Loss of personality"), 1, 0),
    data_collection = ifelse(str_detect(Q3, "Data collection by big tech companies"), 1, 0),
    big_tech    = ifelse(str_detect(Q3, "Dependence on big tech companies"), 1, 0),
    health    = ifelse(str_detect(Q3, "Adverse physical health effects"), 1, 0),
    transparency   = ifelse(str_detect(Q3, "Lack of transparency of how digital products function"), 1, 0),
    skill    = ifelse(str_detect(Q3, "Devaluation of traditional skills"), 1, 0),
    sleep    = ifelse(str_detect(Q3, "Impaired sleep"), 1, 0),
    addiction    = ifelse(str_detect(Q3, "Social media addiction"), 1, 0),
    agency    = ifelse(str_detect(Q3, "Lack of agency"), 1, 0),
    info_overload    = ifelse(str_detect(Q3, "Information overload"), 1, 0),
    alienation    = ifelse(str_detect(Q3, "Alienation"), 1, 0),
    endeavors      = ifelse(str_detect(Q3, "Lack of non-digital endeavors"), 1, 0),
    acceptance = ifelse(Q1 > 3, 1, 0)
  )
```

As a first step, we perform an exploratory analysis to identify which risk categories were selected most frequently.

```{r}
### Transform the dataset
ggplot(df_long, aes(x = reorder(Q3, Q3, FUN = length, decreasing = TRUE))) +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

# ggplot(df_long, aes(x = Q3, fill = factor(acceptance))) +
#   geom_bar(position = "dodge") +
#   labs(x = "Perceived Risk", fill = "Acceptance") +
#   theme_minimal() +
#   theme(axis.text.x = element_text(angle = 90, hjust = 1))

ggplot(df_long, aes(
  x = reorder(Q3, Q3, FUN = length, decreasing = TRUE),
  fill = factor(acceptance)
)) +
  geom_bar(position = "dodge") +
  facet_wrap(~acceptance, ncol = 2) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

In the plot below, dependence on big tech companies is among the most frequently selected responses, ranking second overall, whereas loss of agency is selected less often. When the data is separated by acceptance and non-acceptance of the implant, the overall distributions remain similar, although the ordering of categories changes slightly. Among participants who would accept the implant, dependence on big tech companies represents the most frequently selected concern, whereas among those who would not accept the implant, it ranks second.


As a second step we calculated chi square tests, to see whether people rejecting the implant choose the concerns lack of acency and dependence on big tech companies significantly more often.
```{r}
table(df_test$loss_agency, df_test$acceptance)
chisq.test(df_test$loss_agency, df_test$acceptance)

table(df_test$big_tech, df_test$acceptance)
chisq.test(df_test$big_tech, df_test$acceptance)
```
For loss of agency the difference it not significant. For dependence on bis tech the people who accepted this implant chose this issue more often significantly.

To see if dependence on big tech was chosen significantly more often then the other category, I did a mcnemar test. I did not do the test for lack of agency since we already saw in the plot, that it ranks low. Since I repeat this test several times, I apply bonferroni correction in the end to account for the multiple testing. Dependence on big tech companies is chosen significantly more often then all other reasons except for data collection by big tech companies.

```{r}
mcnemar.test(table(df_test$big_tech, df_test$data_collection))
mcnemar.test(table(df_test$big_tech, df_test$health))
mcnemar.test(table(df_test$big_tech, df_test$loss_personality))
mcnemar.test(table(df_test$big_tech, df_test$transparency))
```


We also analyzed if there was a difference between genders and the age classes, but there was no differencw
```{r}
table(df_test$loss_agency, df_test$gender)
chisq.test(df_test$loss_agency, df_test$gender)

table(df_test$big_tech, df_test$gender)
chisq.test(df_test$big_tech, df_test$gender)
```


```{r}
table(df_test$loss_agency, df_test$age)
chisq.test(df_test$loss_agency, df_test$age)

table(df_test$big_tech, df_test$age)
chisq.test(df_test$big_tech, df_test$age)
```

# Discussion and Conclusion